A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2021-10-24 paper/elsarticle/elsarticle-template.tex
paper

A camera can be calibrated by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}. Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$. Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$. These are obtained using bundle adjustment by minimizing the cumulative reprojection error
\begin{equation}
\sum_{i=1}^{n_\text{pt}} \sum_{j=1}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
\label{re}
\end{equation}
where $\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j \in \mathbb{R}^3$ expresses the point~$\mathbf{X}_{i,j}$ in the local coordinate frame of the camera corresponding to the $j^{\text{th}}$~calibration image and $(\hat{\mathbf{x}}_{i,j}'^\top, 1)^\top \sim \hat{\mathtt{K}}(\hat{\mathtt{R}}_j\mathbf{X}_{i,j} + \hat{\mathbf{t}}_j) \in \mathbb{P}^2$ gives the projection~$\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ in pixels of $\mathbf{X}_{i,j}$ to that image. The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}. Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\hat{\mathtt{R}}_j^{-1}, -\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{}) \in SE(3)$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\hat{\mathtt{R}}_j^{-1}\hat{\mathbf{t}}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.

@@

Given a projector calibration image acquired using the downwards-facing camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(a)); given the scene plane and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:2d}(b)). We obtain the 2D positions of the 2D-3D correspondences by simply running the algorithm for detecting circle centers on the original asymmetric circles pattern image (i.e., the one we feed to the projector). Since---as in the case of recovering chessboard pattern corners---the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain respective 2D-3D correspondences for each projector calibration image, which we use to recover the projector calibration matrix~$\mathtt{K}_\text{proj}$.

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_cam.pdf}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_cam.png}}}
    \caption{Obtaining the virtual camera. It is in accordance with the $X$- and $Y$-axes of the downwards-facing camera the we wish to orient our virtual projector. To avoid projective distortions, the virtual projector is required to face directly downwards to the scene plane; a physical camera placed to face downwards, however, is almost certain to not face downwards 	precisely. Accordingly, we rotate the recovered downwards-facing camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal with respect to the minimum arc-length rotation. It is then the $X$- and $Y$-axes of the \textit{virtual} camera (gray) that we subsequently use to orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera. This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
    \label{fig:virtualcam}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_proj.pdf}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector. The virtual projector is then obtained by (i) rotating the recovered projector (black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Original image.]{{\includegraphics[height=3.4cm]{images/projim.png}}}
    \qquad
    \subfloat[\centering Projecting the original image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray). (a) Original image. (b) Projecting the original image to the scene plane introduces projective distortions if the projector is not fronto-parallel with the scene plane, as is the case here.}
    \label{fig:proj}
\end{figure}

\begin{figure}
    \centering
    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projecting our warped image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray). (a) Original image warped using our approach. (b) After warping the original image according to our approach for the given target location, the image is projected with the effect of having the original image from the viewpoint of the virtual projector. This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and with the desired dimensions in the scene plane, expressed in metric units. Note that pixels shown in black will not be projected to the scene.}
    \label{fig:warp}
\end{figure}

>>>>

A camera can be calibrated by (i) establishing 2D-3D correspondences between pixels in the camera's image plane and the corresponding points in the scene, and on (ii) using those correspondences as input to an optimization procedure that relies on bundle adjustment \cite{triggs1999bundle} to output maximum likelihood estimates of the calibration matrix, the associated lens distortion model coefficients, and, for each calibration image, the pose (i.e., position and orientation) of the camera relative to the 3D scene points from among the 2D-3D correspondences \cite{Hartley2004,zhang2000flexible}. Calibration images containing a calibration surface such as a chessboard pattern of known dimensions and scale are acquired from varying viewpoints using the camera, to be used to establish the 2D-3D correspondences~$\{\mathbf{x}_{i,j} \leftrightarrow \mathbf{X}_{i,j}\}$, $i \in \{ 1, \dots, n_\text{pt} \}, j \in \{ 1, \dots, n_\text{im} \}$, where $n_\text{pt}$ gives the number of correspondences obtained from one calibration image of the pattern, $n_\text{im}$ the number of such images, and $\mathbf{X}_{i,j} = \mathbf{X}_{i,j'}$, for $j, j' \in \{ 1, \dots, n_\text{im} \}$. Bundle adjustment is then used to obtain the maximum likelihood estimates~$(\hat{\omega}, \hat{\mathtt{K}}, \{(\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)\}), j \in \{ 1, \dots, n_\text{im} \}$ of the lens distortion model coefficients~$\omega$, calibration matrix~$\mathtt{K}$, and rigid body transformations~$(\mathtt{R}_j, \mathbf{t}_j) \in SE(3), j \in \{ 1, \dots, n_\text{im} \}$. These are obtained using bundle adjustment by minimizing the cumulative reprojection error
\begin{equation}
\sum_{i=1}^{n_\text{pt}} \sum_{j=1}^{n_\text{im}} d_{\hat{\omega}}(\mathbf{x}_{i,j}, \hat{\mathbf{x}}_{i,j}'),
\label{re}
\end{equation}
where $\hat{\mathbf{x}}_{i,j}' \in \mathbb{R}^2$ gives the projection, in units of pixels, of $\mathbf{X}_{i,j}$ to the image plane of the camera underlying the $j^{\text{th}}$~calibration image, subject to the maximum likelihood estimates $\hat{\mathtt{K}}, (\hat{\mathtt{R}}_j, \hat{\mathbf{t}}_j)$. The function~$d_{\hat{\omega}}$ is a distance function that computes a distance with respect to two pixel positions that it first corrects for lens distortions, in accordance with the maximum likelihood estimate of the lens distortion model coefficients~$\hat{\omega}$ \cite{duane1971close,weng1992camera}. More generally, the projection $\mathbf{x}' \in \mathbb{R}^2$ of $\mathbf{X}$ to the image plane of a camera given calibration matrix~$\mathtt{K}$ and pose~$(\mathtt{R}, \mathbf{t})$ is given by
\begin{equation}
(\mathbf{x}'^\top, 1)^\top \sim \mathtt{K}(\mathtt{R}\mathbf{X} + \mathbf{t}) \in \mathbb{P}^2,
\end{equation}
where $\mathtt{R}\mathbf{X} + \mathbf{t} \in \mathbb{R}^3$ expresses the point~$\mathbf{X}$ in the coordinate frame of the camera. Inverting the rigid body transformation~$(\mathtt{R}_j, \mathbf{t}_j)$ gives the pose~$(\mathtt{R}_j^{-1}, -\mathtt{R}_j^{-1}\mathbf{t}_j^{}) \in SE(3)$ of the $j^{\text{th}}$~camera relative to the 3D scene points of the calibration pattern, with $-\mathtt{R}_j^{-1}\mathbf{t}_j^{} \in \mathbb{R}^3$ the corresponding center of projection.

@@

Given a projector calibration image acquired using the downwards-facing camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(a)); given the scene plane and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection
\begin{equation}
\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2
\end{equation}
of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:2d}(b)). We obtain the 2D positions of the 2D-3D correspondences by simply running the algorithm for detecting circle centers on the original asymmetric circles pattern image (i.e., the one we feed to the projector). Since---as in the case of recovering chessboard pattern corners---the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain respective 2D-3D correspondences for each projector calibration image, which we use to recover the projector calibration matrix~$\mathtt{K}_\text{proj}$.

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_cam.pdf}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_cam.png}}}
    \caption{Obtaining the virtual camera. Our physical downwards-facing camera---in terms of which we wish to orient our augmentations---is almost certain to not face downwards precisely. Accordingly, we rotate the recovered downwards-facing camera (black) about its center of projection such that its optical axis be made parallel with the scene plane's normal with respect to the minimum arc-length rotation. It is then, for each respective target position, the $X$- and $Y$-axes of the resulting \textit{virtual} camera (gray) that we subsequently use to orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera. This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
    \label{fig:virtualcam}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_proj.pdf}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector. For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Original image.]{{\includegraphics[height=3.4cm]{images/projim.png}}}
    \qquad
    \subfloat[\centering Projecting the original image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim.png}}}
    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray) for a given target location. (a) Original image. (b) Projecting the original image to the scene plane introduces projective distortions if the projector is not fronto-parallel with the scene plane, as is the case here.}
    \label{fig:proj}
\end{figure}

\begin{figure}
    \centering
    \subfloat[\centering Our warped image.]{{\includegraphics[height=3.4cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projecting our warped image (downwards view).]{{\includegraphics[height=3.4cm]{images/proj_projim_waped.png}}}
    \caption{Projection to the scene plane from the viewpoint of the recovered projector (bottom left, black; virtual projector in center, gray) for a given target location. (a) Original image warped using our approach. (b) After warping the original image according to our approach for the given target location, the image is projected with the effect of having the original image from the viewpoint of the virtual projector. This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and with the desired dimensions in the scene plane, expressed in metric units. Note that pixels shown in black will not be projected to the scene.}
    \label{fig:warp}
\end{figure}
